### 2013-09-24 evaluation/paper.tex
vlad first mail

\subsection{Superpixel CRFs for Semantic Segmentation}
Our main attention is devoted to the use of conditional random fields for
semantic segmentation.  We use the Pascal VOC and MSRC datasets. A detailed
description of the method will be given in Chapter~\ref{ch:exact_learning}.
%TODO will it?
The Pascal VOC has \ldots training and \ldots validation images, each having
around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
MSRC dataset has \ldots training and \ldots test images, also segmented in
around 100 superpixels each. We use the MSRC-21 dataset, which has 21 target
classes.
Each superpixel is represented as an output variable, with the ground truth obtained
by majority vote over the pixel belonging to the superpixel. We removed all
superpixels which were mostly ``void'', leading to some samples haven much
fewer than 100 variables.
Pairwise potentials were introduced for each pair of neighboring superpixels.
We then removed all superpixels that were labeled as ``void'', together with
their pairwise potentials.
We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC21 dataset.
We use the same three pairwise features for both dataset: a constant, a color
feature and a feature encoding relative vertical position.
Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
2646$ parameters for the potentials on the MSRC21 dataset.

@@

\section{Experiments}
We compare the following algorithms on the above models:
\begin{itemize}
    \item Stochastic Subgradient Descent using the Pegasos schedule: %TODO
    %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
    \item The $1$-slack cutting plane algorithm without constraint caching.
    \item The $1$-slack cutting plane algorithm with constraint caching. For
            each sample, the last $50$ inference results are stored.
            Details of the caching strategy can be found in Chapter~\ref{ch:exact_learning}.
            %TODO
    \item The $n$-slack cutting plane algorithm, where the QP is solved after each training.
    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 training samples.
    \item The SZLJSP algorithm.
\end{itemize}
All algorithm take a $C$ parameter, which we adjusted on a fully trained model
on a validation set (experiments not reported here) and held constant for all models.
We found, however, that the algorithms and models are quite robust to the choice of $C$ within
one or two orders of magnitude.
As stopping criterion, we used a duality gap of $0.1$ when possible, and a
pre-defined number of iterations for the subgradient algorithms. It is worth
noting that the quadratic programming based algorithms have additional
hyper-parameters that we do not discuss here in detail, such as the threshold
for removing a constraint as inactive, how often inactive constraints are
removed, and parameters of the underlying QP solver. On the other hand, the
SZLJSP, for example, has no hyper-parameters except for the stopping criterion.

@@

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction with a tree-structured model. In these
tasks, exact inference is possible, and we can compute the exact objective of
the various algorithms easily. We use the dynamic programming algorithm
implemented in \textsc{OpenGM} for experiments.

@@

Next, we consider the multi-label task, starting with the tree-shaped models.
The results for the yeast and scene datasets are shown in \Figref{yeast_tree} and \Figref{scene_tree},
respectively. The caching $1$-slack algorithm is much faster than the non-caching one on both datasets.
The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
lag behind the stoachastic algorithms on the larger yeast dataset. Again, initial convergence is
faster for the stochastic algorithms.

The results when using the full model are shown in \Figref{yeast_full} and \Figref{scene_full}.
Here, inference is much more complex than in the previous tasks.

>>>>

\subsection{Superpixel CRFs for Semantic Segmentation}
Our main attention is devoted to the use of conditional random fields for
semantic segmentation.  We use the Pascal VOC and MSRC datasets.
The Pascal VOC has 964 training, each having
around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
each. There are 21 classes in the MSRC-21 dataset.
Each superpixel is represented as an output variable, with the ground truth obtained
by majority vote over the pixel belonging to the superpixel. We removed all
superpixels which were mostly ``void'', leading to some samples haven much
fewer than 100 variables.
Pairwise potentials were introduced for each pair of neighboring superpixels.
We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC21 dataset.
We use the same three pairwise features for both dataset: a constant, a color
feature and a feature encoding relative vertical position.
Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
2646$ parameters for the potentials on the MSRC21 dataset.

@@

\section{Experiments}
We compare the following algorithms on the above models:
\begin{itemize}
    \item Stochastic Subgradient Descent (SSGD) using the Pegasos schedule for step-sizes.
    %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
    \item The $1$-slack cutting plane algorithm without constraint caching.
    \item The $1$-slack cutting plane algorithm with constraint caching. For
            each sample, the last $50$ inference results are stored.
            Details of the caching strategy can be found in Chapter~\ref{ch:exact_learning}.
            %TODO
    \item The $n$-slack cutting plane algorithm, where the QP is solved after each training.
    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 training samples.
    \item The SZLJSP algorithm.
\end{itemize}

All algorithm take a $C$ parameter, which we adjusted on a fully trained model
on a validation set (experiments not reported here) and held constant for all
models.  We found, however, that the algorithms and models are quite robust to
the choice of $C$ within one or two orders of magnitude.
As stopping criterion, we used a duality gap of $0.1$ when possible, and a
pre-defined number of iterations for the subgradient algorithms. It is worth
noting that the quadratic programming based algorithms have additional
hyper-parameters that we do not discuss here in detail, such as the threshold
for removing a constraint as inactive, how often inactive constraints are
removed, and parameters of the underlying QP solver. On the other hand, the
SZLJSP, for example, has no hyper-parameters except for the stopping criterion.

@@

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these
tasks, exact inference is possible, and we can compute the exact objective of
the various algorithms easily. We use the dynamic programming algorithm
implemented in \textsc{OpenGM} for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
We found that using the linear programming relaxation on the multi-label
dataset is often tight, and there was no need to resort to approximate
inference.

@@

Next, we consider the multi-label task, starting with the tree-shaped models.
The results for the yeast and scene datasets are shown in \Figref{multi-label-results},
respectively. The caching $1$-slack algorithm is much faster than the non-caching one on both datasets.

The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
lag behind the stoachastic algorithms on the larger yeast dataset. Again, initial convergence is
faster for the stochastic algorithms

%, inference is much more complex than in the previous tasks.

### 2013-09-24 exact_learning/paper.tex
vlad first mail

In this chapter, we investigate the necessity and consequences of these
approximations in the $1$-slack cutting-plane learning algorithm in the context of semantic
image segmentation. We show that despite inference being deemed intractable in
loopy superpixel models, we are able to learn a pairwise conditional random
field model using a structured support vector machine exactly for this task.
We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. Our
approach improves upon the state-of-the art on the MSRC-21 dataset, and is
competative with comparable approaches on the Pascal VOC dataset.% and nyu?

The contribution of this chapter are:
\begin{itemize}
    \item We analyze the simultaneous use of multiple approximate inference
        methods for learning SSVMs using the cutting plane method, relating
        approximate learning to the exact optimum. %FIXME really?
    \item We introduce an efficient caching scheme to accelerate cutting plane
        training.
    \item We demonstrate that using a combination of under-generating and exact
        inference methods, we can learn an SSVM exactly in a practical
        application, even in the presence of loopy graphs.
    \item We show that even using a strong approximate inference procedure can
        improve upon the state-of-the art on the MSRC-21 dataset.
    % also fusion moves is pretty good?
\end{itemize}

@@

\subsection{Implementation Details}
We implemented the described procedure in \pystruct.
We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
the SIFT implementation in the \texttt{vlfeat} package~\citep{vedaldi08vlfeat}.
For clustering visual words, piecewise training of unary potentials and the
approximation to the $\chi^2$-kernel, we made use of the \texttt{scikit-learn}
machine learning package~\citep{pedregosa2011scikit}.
%TODO implementation details here?
The we implement fusion moves with the help of the QPBO-I
method provided by \citet{rother2007optimizing}.
We use the excellent implementation of AD$^3$ provided by
\citet{martins2011augmented}.

>>>>

In this chapter, we investigate the necessity and consequences of these
approximations in the $1$-slack cutting-plane learning algorithm in the context of semantic
image segmentation. We show that despite inference being deemed intractable in
loopy superpixel models, we are able to learn a pairwise conditional random
field model using a structured support vector machine exactly for this task.
We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. Our
approach improves upon the state-of-the-art on the MSRC-21 dataset, and is
competative with comparable approaches on the Pascal VOC dataset.% and nyu?

The contribution of this chapter are:
\begin{itemize}
    \item We analyze the simultaneous use of multiple approximate inference
        methods for learning SSVMs using the cutting plane method, relating
        approximate learning to the exact optimum. %FIXME really?
    \item We introduce an efficient caching scheme to accelerate cutting plane
        training.
    \item We demonstrate that using a combination of under-generating and exact
        inference methods, we can learn an SSVM exactly in a practical
        application, even in the presence of loopy graphs.
    \item We show that even using a strong approximate inference procedure can
        improve upon the state-of-the-art on the MSRC-21 dataset.
    % also fusion moves is pretty good?
\end{itemize}

@@

\subsection{Implementation Details}
We implemented the described procedure in \pystruct.
We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
the SIFT implementation in the \textsc{vlfeat} package~\citep{vedaldi08vlfeat}.
For clustering visual words, piecewise training of unary potentials and the
approximation to the $\chi^2$-kernel, we made use of the \textsc{scikit-learn}
machine learning package~\citep{pedregosa2011scikit}.
%TODO implementation details here?
The we implement fusion moves with the help of the QPBO-I
method provided by \citet{rother2007optimizing}.
We use the excellent implementation of AD$^3$ provided by
\citet{martins2011augmented}.

### 2013-09-24 intro.tex
vlad first mail

In the following we distinguish between the task of semantic segmentation,
which usually distinguishes unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people. We consider four
different datasets in this thesis: the object class segmentation datasets
Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
MSRC-21~\citep{shotton2006textonboost}, and NYU V2~\citep{SilbermanECCV12}.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding,
images in terms of semantic classes. However they employ different mechanisms
to represent and process.

@@

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have received
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown
promising  results for computer vision applications. Using the paradigm of structural
support vector machines, it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, CRFs
allow to combine different cues, possibly produced using different paradgims in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is that context in images is usually represented as a
neighborhood graph of pizels or superpixels.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

@@

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structured_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spatial relations, and outperform
state-of-the-art methods on the NYU V2 datasets.

\section{Publications}
The main material of this thesis has either been published in conference
proceedings or has been submitted to conferences or journals. We now list the
relevant publications.
\begin{description}
    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using
        Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and
        Christoph H. Lampert. Published in the proceedings of the German
        Conference on Pattern Recognition.
    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for
        Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven
        Behnke. Published in the proceedings of the IARP Workshop on Partially
        Supervised Learning.
    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured
        Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to
        the Journal of Machine Learning Research, Open Source Software track.
    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for
        Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke.
        Submitted to the International Conference on Computer Vision Theory and
        Applications.
    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional
        Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven
        Behnke. Submitted to the International Conference on Robotics and
        Automation.
\end{description}

>>>>

In the following we distinguish between the task of semantic segmentation,
which usually distinguishes unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people. We consider four
different datasets in this thesis: the object class segmentation datasets
Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
MSRC-21~\citep{shotton2006textonboost} and NYU V2~\citep{SilbermanECCV12}.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding,
images in terms of semantic classes. However they employ different mechanisms
to represent and process.

@@

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have received
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown
promising  results for computer vision applications. Using the paradigm of structural
support vector machines~(SSVMs), it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, CRFs
allow to combine different cues, possibly produced using different paradgims in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is that context in images is usually represented as a
neighborhood graph of pizels or superpixels.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

@@

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structured_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spatial relations, and outperform
state-of-the-art methods on the NYU V2 datasets.
\pagebreak

\section{Publications}
The main material of this thesis has either been published in conference
proceedings or has been submitted to conferences or journals. We now list the
relevant publications.
\begin{description}
    \item[Chapter~\ref{ch:itm}]%
        \emph{Information Theoretic Clustering using Minimum Spanning Trees}\\
        Andreas C. M\"uller, Sebastian Nowozin and Christoph H. Lampert.
        Published in the proceedings of the German Conference on Pattern
        Recognition.
    \item[Chapter~\ref{ch:semi_supervised}]%
        \emph{Multi-Instance Methods for Partially Supervised Image
        Segmentation}\\
        Andreas C. M\"uller and Sven Behnke. Published in the proceedings of
        the IARP Workshop on Partially Supervised Learning.
    \item[Chapter~\ref{ch:structured_pystruct}]%
        \emph{PyStruct - Structured Prediction in Python}\\
        Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of
        Machine Learning Research, Open Source Software track.
    \item[Chapter~\ref{ch:exact_learning}]%
        \emph{Learning a Loopy Model for Semantic Segmentation Exactly}\\
        Andreas C. M\"uller and Sven Behnke.  Submitted to the International
        Conference on Computer Vision Theory and Applications.
    \item[Chapter~\ref{ch:nyu}]%
        \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}\\
        Andreas C. M\"uller and Sven Behnke. Submitted to the International
        Conference on Robotics and Automation.
\end{description}
